Parameter estimation

### 2023-04-19 SFE-article/Article_Model_Parameters.tex
19.04

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. TO be removed later
\paperwidth=\dimexpr \paperwidth + 6cm\relax
\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{float}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}
\usepackage[bordercolor=gray!20,backgroundcolor=blue!10,linecolor=black,textsize=footnotesize,textwidth=1in]{todonotes}
\setlength{\marginparwidth}{1in}
% \usepackage[utf8]{inputenc}
% \usepackage[nomath]{lmodern}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ are the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient and saturation concentration are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$ and $50^\circ C$ at $200$ bar and $300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.

@@

\section{Introduction}
\subfile{Sections/introduction}

@@

\subsection{Supercritical fluids} \label{CH: Thermodynamic}
\subfile{Sections/Thermo}

@@

\subsection{Governing equations}
The detail derivation of the governing equation can be found in the appendix \ref{CH: Gouverning equations} as well as in the work of \citet{Anderson1995}. Let's assume that any properties of the flow are unifrom across any given cross-section of an channel (or any device like an extractor). Such a flow is called quasi-one-dimensional. The variation of the cross-section might be an result of its irregular shape (or partial filling of an extractor).

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates is given be equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}.

@@

Based on governing equations, the small discontinuity in flow properties, shown on figure \ref{fig: Discontinuity_slow_flow}, can be analysed. The analysis follow the work of \citet{Schreier1982}.

\begin{figure}[h!]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
	\begin{tikzpicture}[]
		\draw (0,2) -- (9,2);	% Top line
		\draw (0,0) -- (9,0); 	% Bottom line
		\draw (4.5,0) -- (4.5,2); 	% Bottom line
		\node at (4,1.5) {$\rho$};
		\node at (5.5,1.5) {$\rho+\delta\rho$};
		\node at (4,1.0) {$P$};
		\node at (5.5,1.0) {$P+\delta P$};
		\node at (4,0.5) {$T$};
		\node at (5.5,0.5) {$T+\delta T$};
		\node at (1,1.0) {$v \rightarrow$};
		\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

The discontinuity is presumed to be at rest relative and \todo[]{Fix the notation of pressure perturbation section to be coherent with governing equations}  the balance equations become

{\footnotesize
	\begin{align*}
		&\rho \delta v + v \delta \rho_f + \delta \rho_f \delta v = 0 \\
		&\delta P = \delta v \delta \rho_f
	\end{align*}
}

@@

If the flow is presumed to be isentropic, $dP/d\rho = c^2$, so $V^2=c^2$, where $c$ is the speed of sound. This can be interpreted that a small pressure wave propagates with the speed of sound relative to the flow.

\subsection{Low Mach number expansion}
\subfile{Sections/Low_mach_number_expansion}

>>>>

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. TO be removed later
\paperwidth=\dimexpr \paperwidth + 6cm\relax
\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{graphicx,float}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}
\usepackage[bordercolor=gray!20,backgroundcolor=blue!10,linecolor=black,textsize=footnotesize,textwidth=1in]{todonotes}
\setlength{\marginparwidth}{1in}
% \usepackage[utf8]{inputenc}
% \usepackage[nomath]{lmodern}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ are the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
We consider a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ denotes the state variables, $\Theta$ represents the parameters, and $u$ is the control variables. Our focus is on the supercritical extraction process, which involves a partially filled extractor with a fixed bed that operates under constant conditions. We assume the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. To model this two-dimensional case, we use the concept of quasi-one-dimensional flow.
To describe the fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent, we use a distributed-parameter model based on \citet{Reverchon1996}. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and saturation concentration. We estimate these parameters from four experiments at different temperatures and pressures: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. We obtain the parameter values using the maximum likelihood estimation method on yield data under the normal error assumption.

@@

\section{Introduction}
%\subfile{Sections/introduction}
\subfile{Sections/introduction_imp}

@@

\subsection{Supercritical fluids} \label{CH: Thermodynamic}
%\subfile{Sections/Thermo}
\subfile{Sections/Thermo_imp}

@@

\subsection{Governing equations}
The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}.  Quasi-one-dimensional flow is a type of fluid flow that is characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as a irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates are given by equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}.

@@

Based on governing equations, the small discontinuity (defined as $\delta$) in flow properties, shown on figure \ref{fig: Discontinuity_slow_flow}, can be analysed. The analysis follow the work of \citet{Schreier1982}.

\begin{figure}[!ht]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {$\rho_f$};
			\node at (5.5,1.5) {$\rho_f+\delta\rho_f$};
			\node at (4,1.0) {$P$};
			\node at (5.5,1.0) {$P+\delta P$};
			\node at (4,0.5) {$T$};
			\node at (5.5,0.5) {$T+\delta T$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

The discontinuity is presumed to be at rest relative and the balance equations become

{\footnotesize
	\begin{align*}
		&\rho_f \delta v + v \delta \rho_f + \delta \rho_f \delta v = 0 \\
		&\delta P = \delta v \delta \rho_f
	\end{align*}
}

@@

If the flow is presumed to be isentropic, $dP/d\rho = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted that a small pressure wave propagates with the speed of sound relative to the flow.

\subsection{Low Mach number expansion}
%\subfile{Sections/Low_mach_number_expansion}
\subfile{Sections/Low_mach_number_expansion_imp}

### 2023-04-19 SFE-article/Sections/Low_mach_number_expansion_imp.tex
19.04

	As discussed by \citet{Lions2013}, the low Mach number equations are a subset of the fully compressible equations of motion (continuity, momentum and energy). Such a set of equations allow for large variations in gas density but it is consider to be acoustically incompressible. The low Mach number equations are preferred over the full compressible equations for low speed flow problems $\left( M_a = \frac{|V|}{\sqrt{ \partial P / \partial \rho }} \ll 1 \right)$ to avoids the need to resolve fast-moving acoustic signals. The equations are derived from the compressible equations based on the perturbation theory. The perturbation theory develops an expression for the desired solution in terms of a formal power series known as a perturbation series in some "small" parameter $\zeta$, that quantifies the deviation from the exactly solvable problem. The leading term in this power series is the solution of the exactly solvable problem, while further terms describe the deviation in the solution, due to the deviation from the initial problem.

	The equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_2} describe the fully compressible equations of motion (respectively the transport of mass, momentum and energy) for quasi-one-dimensional case. We rescale the time variable, considering finally

@@

	The conservative non-dimensional equations of motion becomes

@@

	Let's define $\zeta=M_a^2$ and assume small Mach numbers, $M_a \ll 1$, then the kinetic energy, viscous work, and gravity work terms can be neglected in the energy equation since those terms are scaled by the square of the Mach number. The inverse of Mach number squared remains in the momentum equations, suggesting singular behaviour. In order to explore the singularity, the pressure, velocity and temperature are expanded as asymptotic series in terms of the parameter $\zeta$

@@

	 By expanding performing power expansion on the continuity equation and taking the limit of $\zeta$ from the positive side we get

@@

	The continuity equation become

@@

	The form of the continuity equation stays the same. Considering the momentum equation, it can be seen that the inverse of Mach number squared remains which suggests singular behavior.

@@

	The first two terms stays the same, but the third one become different in structure. By further investigation of the pressure term in the momentum equation it can be observed

@@

	The simplification of the $P_0$ in the momentum equation comes from the fact that $P_0$ is independent of $z$. As it was presented above, the thermodynamical pressure moves with speed of sound, and any perturbation propagates instantaneously. The term related to $P_2$ and higher order terms become zero at the limit of $M_a \rightarrow 0$. The momentum equation become

@@

	By expanding performing power expansion on the energy equation and taking the limit of $\zeta$ from the positive side we get

@@

	The non-conservative form of the energy equation become

@@

	If the calorically perfect gas is assumed then $e=C_vT$, where $C_v$ is the constant specific heat. The energy equation can derived in terms of temperature $T$

@@

	If isothermal case is assumed then, the energy equation becomes

@@

		A general formulation the internal energy for a real gas is:

@@

	The internal energy is a function of two intensive properties, in this case $T$ and $v_m=1/\rho_f$. But, in the case of an ideal gas, the equation of state is such that the second term in this equation is identically equal to zero. So the ideal gas is a special case in which the molar internal energy is a function only of temperature. For Peng-Robinson equation of state, the internal energy is defined as

@@

	\todo[]{Add REF}

@@

	It can be deduce that the continuity equation becomes $\frac{\partial \rho_f}{\partial t} = 0$ at constant temperature and pressure. Moreover, the incompressibility condition $div(A_fV)=0$ is obtained.

	Assuming an arbitrary function $\hat{P}$, which describe the total pressure, the dimensional momentum equation can be written as

@@

	From the incompressibility conditions we can deduce that

@@

	The final form of the momentum equation corresponds to the Bernoulli's principle

@@

	The Bernoulli's principle can be used to find the hydrodynamic pressure caused by varying cross-section at steady-state. Moreover, if the flow velocity is relatively low and so all pressure changes are hydrodynamic (due to velocity motion) rather than thermodynamic. The effect of this is that $\partial \rho / \partial P = 0$. In other words, the small changes in pressure due to flow velocity changes do not change the density. This has a secondary effect -- the speed of sound in the fluid is $\partial P/\partial \rho = \infty$	in this instance. So there is an infinite speed of sound, which makes the equations elliptic in nature. It can be deduced that at the isothermal conditions the density in the system propagates with the same speed as pressure due to the fact that they are both connected through the equation of state.

>>>>

	As discussed by \citet{Lions2013}, the low Mach number equations are a subset of the fully compressible equations of motion (continuity, momentum and energy). Such a set of equations allow for large variations in gas density, but it is considered acoustically incompressible. Therefore, the low Mach number equations are preferred over the full compressible equations for low-speed flow problems $\left( M_a = \frac{|v|}{\sqrt{ \partial P / \partial \rho_f }} \ll 1 \right)$ to avoids the need to resolve fast-moving acoustic signals. The equations are derived from compressible equations based on the perturbation theory. The perturbation theory develops an expression for the desired solution in terms of a formal power series known as a perturbation series in some "small" parameter $\zeta$, that quantifies the deviation from the exactly solvable problem. The leading term in this power series is the solution of the exactly solvable problem, while further terms describe the deviation in the solution due to the deviation from the initial problem.

	The equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_2} describe the fully compressible equations of motion (respectively, the transport of mass, momentum and energy) for the quasi-one-dimensional case. We rescale the time variable, considering finally

@@

	The conservative non-dimensional equations of motion become

@@

	Let's define $\zeta=M_a^2$ and assume small Mach numbers, $M_a \ll 1$, then the kinetic energy, viscous work, and gravity work terms can be neglected in the energy equation since the square of the Mach number scales those terms. The inverse of Mach number squared remains in the momentum equations, suggesting singular behaviour. In order to explore the singularity, the pressure, velocity and temperature are expanded as asymptotic series in terms of the parameter $\zeta$

@@

	By expanding performing power expansion on the continuity equation and taking the limit of $\zeta$ from the positive side, we get

@@

	The continuity equation becomes

@@

	The form of the continuity equation stays the same. Considering the momentum equation, it can be seen that the inverse of Mach number squared remains, which suggests singular behaviour.

@@

	The first two terms stay the same, but the third one becomes different in structure. By further investigation of the pressure term in the momentum equation, it can be observed.

@@

	The simplification of the $P_0$ in the momentum equation comes from the fact that $P_0$ is independent of $z$. As was presented above, the thermodynamical pressure moves with the speed of sound, and any perturbation propagates instantaneously. The term related to $P_2$ and higher order terms become zero at the limit of $M_a \rightarrow 0$. The momentum equation becomes

@@

	By expanding performing power expansion on the energy equation and taking the limit of $\zeta$ from the positive side, we get

@@

	The non-conservative form of the energy equation becomes

@@

	If the calorically perfect gas is assumed, then $e=C_vT$, where $C_v$ is the constant specific heat. The energy equation can be derived in terms of temperature $T$.

@@

	If an isothermal case is assumed, then the energy equation becomes

@@

	As presented by \citet{Elliott2011}, a general formulation of the internal energy for a real gas is:

@@

	The internal energy is a function of two intensive properties, in this case, $T$ and $v_m=1/\rho_f$. But, in the case of an ideal gas, the equation of state is such that the second term in this equation is identically equal to zero. So the ideal gas is a special case in which the molar internal energy is a function only of temperature. For Peng-Robinson equation of state, the internal energy is defined as

@@

@@

	It can be deduced that the continuity equation becomes $\frac{\partial \rho_f}{\partial t} = 0$ at constant temperature and pressure. Moreover, the incompressibility condition $div(A_fV)=0$ is obtained.

	Assuming an arbitrary function $\hat{P}$, which describes the total pressure, the dimensional momentum equation can be written as

@@

	From the incompressibility conditions, we can deduce that

@@

	The final form of the momentum equation corresponds to Bernoulli's principle

@@

	Bernoulli's principle can be used to find the hydrodynamic pressure caused by varying cross-sections at steady-state. Moreover, if the flow velocity is relatively low, all pressure changes are hydrodynamic (due to velocity motion) rather than thermodynamic. The effect of this is that $\partial \rho / \partial P = 0$. In other words, the small changes in pressure due to flow velocity changes do not change the density. This has a secondary effect -- the speed of sound in the fluid is $\partial P/\partial \rho = \infty$ in this instance. So there is an infinite speed of sound, which makes the equations elliptic in nature. It can be deduced that at the isothermal conditions, the density in the system propagates with the same speed as pressure since they are both connected through the equation of state.
